\LARGE \bf Compressed Directional Distance Transform for Fast 2D Ray Casting

### 2017-02-25 paper/main.tex
layed out figures nicely

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphicx}
\usepackage{tabu}
% \usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

@@

\begin{figure}[h]
\img{synthetic_cropped.png}
\caption{Occupancy grid map for the synthetic map example.}
\label{violin:basement:grid}
\end{figure}

@@

@@

\begin{figure}[h]
\img{grid_trial2_basement_cropped.png}
\caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method. Basement map. X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
\label{violin:basement:grid}
\end{figure}

\begin{figure}[h]
\img{grid_trial2_synthetic_cropped.png}
\caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method. Synthetic map. X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
\label{violin:basement:grid}
\end{figure}

\begin{figure}[h]
\img{random_trial2_basement_cropped.png}
\caption{Violin plots for a large number of randomly sampled queries. Basement map. The random access pattern degrades the caching performance variably for each method.}
\label{violin:basement:random}
\end{figure}

\begin{figure}[h]
\img{random_trial2_synthetic_cropped.png}
\caption{Violin plots for a large number of randomly sampled queries. Synthetic map. The random access pattern degrades the caching performance variably for each method.}
\label{violin:basement:random}
\end{figure}

@@

\begin{figure}[h]
\begin{center}
\begin{tabular}{ | m{8em} | m{2.3cm}| m{1.5cm} | } 
\hline
 \multicolumn{3}{|c|}{Basement Map, $\theta$ discretization: 108} \\
 \hline
 Method & Memory Usage & Init. Time \\
 \hline
 Bresenham's Line & 1.37 MB & 0.006 sec  \\
 Ray Marching & 5.49 MB & 0.16 sec  \\
 CDDT & 6.34 MB & 0.067 sec  \\
 PCDDT & 4.07 MB & 2.2 sec  \\
 Lookup Table & 296.63 MB & 15.3 sec  \\
\hline
\end{tabular}
\end{center}
\caption{The construction time and memory footprint of each method for the Basement map.}
\label{table:basement:init}
\end{figure}

\begin{figure}[h]
\begin{center}
\begin{tabular}{ | m{8em} | m{2.3cm}| m{1.5cm} | } 
\hline
 \multicolumn{3}{|c|}{Synthetic Map, $\theta$ discretization: 108} \\
 \hline
 Method & Memory Usage & Init. Time \\
 \hline
 Bresenham's Line & 1 MB & 0.004 sec  \\
 Ray Marching  & 4 MB & 0.13 sec  \\
 CDDT & 2.71 MB & 0.03 sec  \\
 PCDDT & 1.66 MB & 0.74 sec  \\
 Lookup Table & 216 MB & 9.1 sec  \\
\hline
\end{tabular}
\end{center}
\caption{The construction time and memory footprint of each method for the Basement map.}
\label{table:basement:init}
\end{figure}

@@

This work demonstrates that the proposed CDDT algorithm may be used in mobile robotics to accelerate sensor model computation when localizing in a two dimensional occupancy grid map. While the precomputed LUT approach is generally 1.1 to 1.7 times faster than the proposed algorithm, the memory footprint of the proposed data structure is significantly smaller which may be desirable for resource constrained systems. Other than the LUT approach, CDDT is significantly faster than the other methods evaluated. Ray marching is the next best, ranging from 1.28 to 2.25 times slower than PCDDT depending on access patterns and environmental factors. The comparison with the widely used Bresenham's Line algorithm is more stark, ranging from a factor of 5.39 to 14.8 speedup in our benchmarks.

@@

\begin{figure}[h]
\begin{center}
\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
\hline
\multicolumn{5}{|c|}{Basement Map, Random Sampling Ray Cast} \\
\hline
Method & Mean & Median & IQR & Speedup \\
\hline
BL & 9.66e-07 & 1.05e-06 & 1.08e-06 & 1 \\
RM & 2.12e-07 & 1.68e-07 & 1.64e-07 & 4.56 \\
CDDT & 1.58e-07 & 1.49e-07 & 9.1e-08 & 6.13 \\
PCDDT & 1.41e-07 & 1.32e-07 & 6.5e-08 & 6.83 \\
LUT & 1.89e-07 & 1.7e-07 & 2.1e-08 & 5.10 \\
\hline
\end{tabular}
\end{center}
\caption{The construction time and memory footprint of each method for the Basement map.}
\label{table:basement:init}
\end{figure}

\begin{figure}[h]
\begin{center}
\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
\hline
\multicolumn{5}{|c|}{Basement Map, Grid Sampling Ray Cast} \\
\hline
Method & Mean & Median & IQR & Speedup \\
\hline
BL & 8.29e-07 & 9.24e-07 & 9.53e-07 & 1 \\
RM & 1.65e-07 & 1.26e-07 & 1.34e-07 & 5.02 \\
CDDT & 7.69e-08 & 7.2e-08 & 1.6e-08 & 10.78 \\
PCDDT & 7.32e-08 & 7e-08 & 1.4e-08 & 11.33 \\
LUT & 6.13e-08 & 4.3e-08 & 4.6e-08 & 13.53 \\
\hline
\end{tabular}
\end{center}
\caption{The construction time and memory footprint of each method for the Basement map.}
\label{table:basement:init}
\end{figure}

\begin{figure}[h]
\begin{center}
\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
\hline
\multicolumn{5}{|c|}{Basement Map, Particle Filter Ray Cast} \\
\hline
Method & Mean & Median & IQR & Speedup \\
\hline
BL & 3.89e-07 & 2.67e-07 & 2.31e-07 & 1 \\
RM & 1.56e-07 & 1.18e-07 & 9.4e-08 & 2.49 \\
CDDT & 7.67e-08 & 7.8e-08 & 5.7e-08 & 5.07 \\
PCDDT & 7.22e-08 & 7.4e-08 & 5.4e-08 & 5.39 \\
LUT & 4.35e-08 & 4.1e-08 & 5e-09 & 8.94 \\
\hline
\end{tabular}
\end{center}
\caption{The construction time and memory footprint of each method for the Basement map.}
\label{table:basement:init}
\end{figure}

\begin{figure}[h]
\begin{center}
\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
\hline
\multicolumn{5}{|c|}{Synthetic Map, Random Sampling Ray Cast} \\
\hline
Method & Mean & Median & IQR & Speedup \\
\hline
BL & 1.19e-06 & 1.41e-06 & 7.71e-07 & 1 \\
RM & 1.52e-07 & 1.25e-07 & 1.05e-07 & 7.81 \\
CDDT & 1.24e-07 & 1.05e-07 & 5.3e-08 & 9.59 \\
PCDDT & 1.19e-07 & 1.01e-07 & 5e-08 & 10.02 \\ 
LUT & 1.82e-07 & 1.68e-07 & 1.4e-08 & 6.55 \\
\hline
\end{tabular}
\end{center}
\caption{The construction time and memory footprint of each method for the Basement map.}
\label{table:basement:init}
\end{figure}

\begin{figure}[h]
\begin{center}
\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
\hline
\multicolumn{5}{|c|}{Synthetic Map, Grid Sampling Ray Cast} \\
\hline
Method & Mean & Median & IQR & Speedup \\
\hline
BL & 1.03e-06 & 1.20e-06 & 6.79e-07 & 1 \\ 
RM & 1.27e-07 & 1.03e-07 & 1.06e-07 & 8.06 \\
CDDT & 7.02e-08 & 6.8e-08 & 1e-08 & 14.63 \\
PCDDT & 6.94e-08 & 6.8e-08 & 9e-09 & 14.80 \\
LUT & 6.33e-08 & 4.2e-08 & 4.6e-08 & 16.21 \\
\hline
\end{tabular}
\end{center}
\caption{The construction time and memory footprint of each method for the Basement map.}
\label{table:basement:init}
\end{figure}

@@

\end{thebibliography}
\end{document}

>>>>

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphicx}
\usepackage{tabu}
\usepackage{float}
% \usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

@@

@@

% \begin{figure}[h]
% \img{grid_trial2_basement_cropped.png}
% \caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method. Basement map. X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
% \label{violin:basement:grid}
% \end{figure}

\begin{figure}[ht]
\img{basement_grid_random_cropped.png}
% \begin{center}\includegraphics[width=248px,height=520px]{basement_benchmarks_cropped.png}\end{center}
\caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method. Basement map. X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
\label{violin:basement:all}
\end{figure}

@@

% \begin{figure}[h]
% \img{grid_trial2_basement_cropped.png}
% \caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method. Basement map. X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
% \label{violin:basement:grid}
% \end{figure}

% \begin{figure}[h]
% \img{grid_trial2_synthetic_cropped.png}
% \caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method. Synthetic map. X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
% \label{violin:basement:grid}
% \end{figure}

% \begin{figure}[h]
% \img{random_trial2_basement_cropped.png}
% \caption{Violin plots for a large number of randomly sampled queries. Basement map. The random access pattern degrades the caching performance variably for each method.}
% \label{violin:basement:random}
% \end{figure}

% \begin{figure}[h]
% \img{random_trial2_synthetic_cropped.png}
% \caption{Violin plots for a large number of randomly sampled queries. Synthetic map. The random access pattern degrades the caching performance variably for each method.}
% \label{violin:basement:random}
% \end{figure}

@@

\begin{figure}[ht]
\begin{center}
\begin{tabular}{ | m{8em} | m{2.3cm}| m{1.5cm} | } 
\hline
 \multicolumn{3}{|c|}{Basement Map, $\theta$ discretization: 108} \\
 \hline
 Method & Memory Usage & Init. Time \\
 \hline
 Bresenham's Line & 1.37 MB & 0.006 sec  \\
 Ray Marching & 5.49 MB & 0.16 sec  \\
 CDDT & 6.34 MB & 0.067 sec  \\
 PCDDT & 4.07 MB & 2.2 sec  \\
 Lookup Table & 296.63 MB & 15.3 sec  \\
\hline
\end{tabular}
\end{center}
\caption{The construction time and memory footprint of each method for the Basement map.}
\label{table:basement:init}
\end{figure}

\begin{figure}[ht]
\begin{center}
\begin{tabular}{ | m{8em} | m{2.3cm}| m{1.5cm} | } 
\hline
 \multicolumn{3}{|c|}{Synthetic Map, $\theta$ discretization: 108} \\
 \hline
 Method & Memory Usage & Init. Time \\
 \hline
 Bresenham's Line & 1 MB & 0.004 sec  \\
 Ray Marching  & 4 MB & 0.13 sec  \\
 CDDT & 2.71 MB & 0.03 sec  \\
 PCDDT & 1.66 MB & 0.74 sec  \\
 Lookup Table & 216 MB & 9.1 sec  \\
\hline
\end{tabular}
\end{center}
\caption{The construction time and memory footprint of each method for the Synthetic map.}
\label{table:synthetic:init}
\end{figure}

@@

This work demonstrates that the proposed CDDT algorithm may be used in mobile robotics to accelerate sensor model computation when localizing in a two dimensional occupancy grid map. While the precomputed LUT approach is generally 1.1 to 1.7 times faster than the proposed algorithm, the memory footprint of the proposed data structure is significantly smaller which may be desirable for resource constrained systems. Other than the LUT approach, CDDT is significantly faster than the other methods evaluated. Ray marching is the next best, ranging from 1.28 to 2.25 times slower than PCDDT depending on access patterns and environmental factors. The comparison with the widely used Bresenham's Line algorithm is more stark, ranging from a factor of 5.39 to 14.8 in our benchmarks.

@@

\begin{figure}[h!]
\img{synthetic_cropped.png}
\caption{Synthetic map occupancy grid. 1024x1024}
\label{violin:basement:grid}
\end{figure}

\begin{figure}[h!]
\begin{center}
\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}
\begin{tabular}{ | C{1.15cm} | C{1.25cm}| C{1.25cm} | C{1.25cm} | C{1.25cm}|  } 
\hline
\multicolumn{5}{|c|}{\textbf{Synthetic Map Ray Cast Benchmarks}} \\
\hline \hline
\multicolumn{5}{|c|}{Random Sampling} \\
\hline
Method & Mean & Median & IQR & Speedup \\
\hline
BL & 1.19e-06 & 1.41e-06 & 7.71e-07 & 1 \\
RM & 1.52e-07 & 1.25e-07 & 1.05e-07 & 7.81 \\
CDDT & 1.24e-07 & 1.05e-07 & 5.3e-08 & 9.59 \\
PCDDT & 1.19e-07 & 1.01e-07 & 5e-08 & 10.02 \\ 
LUT & 1.82e-07 & 1.68e-07 & 1.4e-08 & 6.55 \\
\hline \hline
\multicolumn{5}{|c|}{Grid Sampling} \\
\hline
Method & Mean & Median & IQR & Speedup \\
\hline
BL & 1.03e-06 & 1.20e-06 & 6.79e-07 & 1 \\ 
RM & 1.27e-07 & 1.03e-07 & 1.06e-07 & 8.06 \\
CDDT & 7.02e-08 & 6.8e-08 & 1e-08 & 14.63 \\
PCDDT & 6.94e-08 & 6.8e-08 & 9e-09 & 14.80 \\
LUT & 6.33e-08 & 4.2e-08 & 4.6e-08 & 16.21 \\
\hline
\end{tabular}
\caption{All times listed in seconds, speedup relative to Bresenham's Line}
\label{table:basement:init}
\end{center}
\end{figure}

\begin{figure}[h!]
\begin{center}
\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}
\begin{tabular}{ | C{1.15cm} | C{1.25cm}| C{1.25cm} | C{1.25cm} | C{1.25cm}|  } 
\hline
\multicolumn{5}{|c|}{\textbf{Basement Map Ray Cast Benchmarks}} \\
\hline \hline
Method & Mean & Median & IQR & Speedup \\
\hline
\multicolumn{5}{|c|}{Random Sampling} \\
\hline
BL & 9.66e-07 & 1.05e-06 & 1.08e-06 & 1 \\
RM & 2.12e-07 & 1.68e-07 & 1.64e-07 & 4.56 \\
CDDT & 1.58e-07 & 1.49e-07 & 9.1e-08 & 6.13 \\
PCDDT & 1.41e-07 & 1.32e-07 & 6.5e-08 & 6.83 \\
LUT & 1.89e-07 & 1.7e-07 & 2.1e-08 & 5.10 \\
\hline \hline
\multicolumn{5}{|c|}{Grid Sampling} \\
\hline
Method & Mean & Median & IQR & Speedup \\
\hline
BL & 8.29e-07 & 9.24e-07 & 9.53e-07 & 1 \\
RM & 1.65e-07 & 1.26e-07 & 1.34e-07 & 5.02 \\
CDDT & 7.69e-08 & 7.2e-08 & 1.6e-08 & 10.78 \\
PCDDT & 7.32e-08 & 7e-08 & 1.4e-08 & 11.33 \\
LUT & 6.13e-08 & 4.3e-08 & 4.6e-08 & 13.53 \\
\hline \hline
\multicolumn{5}{|c|}{Particle Filter} \\
\hline
Method & Mean & Median & IQR & Speedup \\
\hline
BL & 3.89e-07 & 2.67e-07 & 2.31e-07 & 1 \\
RM & 1.56e-07 & 1.18e-07 & 9.4e-08 & 2.49 \\
CDDT & 7.67e-08 & 7.8e-08 & 5.7e-08 & 5.07 \\
PCDDT & 7.22e-08 & 7.4e-08 & 5.4e-08 & 5.39 \\
LUT & 4.35e-08 & 4.1e-08 & 5e-09 & 8.94 \\
\hline
\end{tabular}

\caption{All times listed in seconds, speedup relative to Bresenham's Line}
\label{table:basement:init}
\end{center}
\end{figure}

% \begin{figure}[h!]
% \begin{center}
% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
% \hline
% \multicolumn{5}{|c|}{Basement Map, Random Sampling Ray Cast} \\
% \hline
% Method & Mean & Median & IQR & Speedup \\
% \hline
% BL & 9.66e-07 & 1.05e-06 & 1.08e-06 & 1 \\
% RM & 2.12e-07 & 1.68e-07 & 1.64e-07 & 4.56 \\
% CDDT & 1.58e-07 & 1.49e-07 & 9.1e-08 & 6.13 \\
% PCDDT & 1.41e-07 & 1.32e-07 & 6.5e-08 & 6.83 \\
% LUT & 1.89e-07 & 1.7e-07 & 2.1e-08 & 5.10 \\
% \hline
% \end{tabular}
% \end{center}
% \caption{Statistics per ray cast query for random sampling synthetic map.}
% \label{table:basement:init}
% \end{figure}
% \begin{figure}[h!]
% \begin{center}
% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
% \hline
% \multicolumn{5}{|c|}{Basement Map, Grid Sampling Ray Cast} \\
% \hline
% Method & Mean & Median & IQR & Speedup \\
% \hline
% BL & 8.29e-07 & 9.24e-07 & 9.53e-07 & 1 \\
% RM & 1.65e-07 & 1.26e-07 & 1.34e-07 & 5.02 \\
% CDDT & 7.69e-08 & 7.2e-08 & 1.6e-08 & 10.78 \\
% PCDDT & 7.32e-08 & 7e-08 & 1.4e-08 & 11.33 \\
% LUT & 6.13e-08 & 4.3e-08 & 4.6e-08 & 13.53 \\
% \hline
% \end{tabular}
% \end{center}
% \caption{Statistics per ray cast query for grid sampling synthetic map.}
% \label{table:basement:init}
% \end{figure}
% \begin{figure}[h!]
% \begin{center}
% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.15cm}|  } 
% \hline
% \multicolumn{5}{|c|}{Basement Map, Particle Filter Ray Cast} \\
% \hline
% Method & Mean & Median & IQR & Speedup \\
% \hline
% BL & 3.89e-07 & 2.67e-07 & 2.31e-07 & 1 \\
% RM & 1.56e-07 & 1.18e-07 & 9.4e-08 & 2.49 \\
% CDDT & 7.67e-08 & 7.8e-08 & 5.7e-08 & 5.07 \\
% PCDDT & 7.22e-08 & 7.4e-08 & 5.4e-08 & 5.39 \\
% LUT & 4.35e-08 & 4.1e-08 & 5e-09 & 8.94 \\
% \hline
% \end{tabular}
% \end{center}
% \caption{Statistics per ray cast query for random sampling basement map.}
% \label{table:basement:init}
% \end{figure}
% \begin{figure}[h!]
% \begin{center}
% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.15cm}|  } 
% \hline
% \multicolumn{5}{|c|}{Synthetic Map, Random Sampling Ray Cast} \\
% \hline
% Method & Mean & Median & IQR & Speedup \\
% \hline
% BL & 1.19e-06 & 1.41e-06 & 7.71e-07 & 1 \\
% RM & 1.52e-07 & 1.25e-07 & 1.05e-07 & 7.81 \\
% CDDT & 1.24e-07 & 1.05e-07 & 5.3e-08 & 9.59 \\
% PCDDT & 1.19e-07 & 1.01e-07 & 5e-08 & 10.02 \\ 
% LUT & 1.82e-07 & 1.68e-07 & 1.4e-08 & 6.55 \\
% \hline
% \end{tabular}
% \end{center}
% \caption{The construction time and memory footprint of each method for the Basement map.}
% \label{table:basement:init}
% \end{figure}
% \begin{figure}[h!]
% \begin{center}
% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.15cm}|  } 
% \hline
% \multicolumn{5}{|c|}{Synthetic Map, Grid Sampling Ray Cast} \\
% \hline
% Method & Mean & Median & IQR & Speedup \\
% \hline
% BL & 1.03e-06 & 1.20e-06 & 6.79e-07 & 1 \\ 
% RM & 1.27e-07 & 1.03e-07 & 1.06e-07 & 8.06 \\
% CDDT & 7.02e-08 & 6.8e-08 & 1e-08 & 14.63 \\
% PCDDT & 6.94e-08 & 6.8e-08 & 9e-09 & 14.80 \\
% LUT & 6.33e-08 & 4.2e-08 & 4.6e-08 & 16.21 \\
% \hline
% \end{tabular}
% \end{center}
% \caption{The construction time and memory footprint of each method for the Basement map.}
% \label{table:basement:init}
% \end{figure}

@@

\end{thebibliography}

\end{document}
